\section*{Conclusion}
In this thesis, we have shown properties of transducer degrees. We showed the concept of function streams as a way to create new streams. Furthermore, we gave the definition of spiralling functions and some of their properties. We gave the definition of zero loops and introduced block loops. We used these loops to give a meaningful structure to the transducts of spiralling function streams. Moreover, we have shown definitions concerning weights and masses and used this to compress the transducts of spiralling function streams to a new form. We used this form to prove that the degree of $\fs{n^2}$ is an atom. We briefly sketched why not all sets of degrees have a supremum. We list the properties that we touched upon in this thesis below.

\section*{Some properties}
\begin{itemize}
	\item There exists a bottom degree, namely, the degree of all ultimately periodic sequences.
	\item Every countable set of degrees has an upper bound.
	\item The degree of $\fs{n}$ is an atom.
	\item The degree of $\fs{n^2}$ is an atom.
	\item The degree of $\fs{n^k}$ is not atom for $k\geq 3$.
	\item There are countably many atoms.
	\item The set of of degrees $S = \{[\fs{2^{2^n}}],[\fs{3^{3^{3^n}}}]\}$ does not have a supremum.
\end{itemize}

\section*{Future Work}
There are still a lot of open questions around the topic of transducer degrees. Some that are connected to the concepts explored in this thesis are mentioned below:
\begin{enumerate}
	\item Is the degree of the Thue-Morse stream an atom?
	\item Are there uncountably many atoms?
	\item Are there degrees that only transduce to atom degrees or the zero degree?
	\item Are there sets of degrees that do have a supremum. (We have seen that this is not the case in general).
\end{enumerate}
The research concerning transducer degrees is still very active. The proof that there are degrees without a supremum was only published online in december of 2019. We recommend to keep an eye on the website of Jörg Endrullis to be informed about the latest discoveries in this field \cite{websiteofjeorg}.
